Comparison of Dimensionality Reduction Schemes for Parallel Global Optimization Algorithms

### 2018-09-16 2018/RSD/MultiEvolvents/slides/slides.tex
RSD: minir changes in slides
RSD: update slides

\setromanfont{CMU Serif}
%\setsansfont{CMU Sans Serif}
\setmathfont{Latin Modern Math}

@@

\begin{frame}
  \frametitle{Basic parallel optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x^k)\) and the function \(\varphi(y(x^k))\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1}=d(t)\in (x_{t-1}, x_t)
    \end{displaymath}
    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

\begin{frame}
  \frametitle{Parallel optimization method with multiple evolvents}

@@

\begin{frame}
  \frametitle{Results of comparison}
  \begin{figure}[ht]
    \hspace*{-0.9cm}
    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_opt_pt_op.pdf} }}
    \subfloat[\(r=5.0\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }} \hspace*{4cm}
    \caption{Operating characteristics on GKLS 2d Simple class}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Results of comparison}
  \begin{figure}[ht]
    \hspace*{-0.9cm}
    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
    \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_same_r_opt_pt_op.pdf} }}
    \caption{Operating characteristics on GKLS 3d Simple class}
  \end{figure}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusions}
    %Already done:
    \begin{itemize}
      \item the smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
      \item the shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint. About 95\% of iterations are overhead to fight the auxiliary constraint.
      \item rotated evolvents perform almost the same as the shifted ones but without overhead.
    \end{itemize}
    %Future work:
    %\begin{itemize}
    %  \item
    %\end{itemize}
\end{frame}

\begin{frame}{{}}
  \frametitle{ }
  \begin{center}
    \Large{Q\&A}

\vspace{1cm}

>>>>

\setromanfont{CMU Serif}
%\setsansfont{CMU Sans Serif}
\setmathfont{Latin Modern Math}
\setlength{\tabcolsep}{1pt}

@@

\begin{frame}
  \frametitle{Basic parallel optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x^{k+j})\) and the function \(\varphi(y(x^{k+j}))\), \(j=\overline{1,p}\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose \(p\) intervals \((x_{t_j-1}, x_{t_j})\) with the greatest characteristics and
    compute objective \(f(y(x^{k+j}))\) in points chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1+j}=d(t)\in (x_{t_j-1}, x_{t_j}),\:j=\overline{1,p}
    \end{displaymath}
    \item[Step 5.] If \(x_{t_j}-x_{t_j-1}<\varepsilon\) for one of \(j=\overline{1,p}\), stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

\begin{frame}
  \frametitle{Parallel optimization method with multiple evolvents}
  Using the multiple mapping allows solving initial problem by parallel solving the problems
  \[
  \min\{\varphi(y^s(x)):x\in [0,1]\}, 1\leqslant s\leqslant S
  \]
  on a set of intervals $[0,1]$ by the index method. Each one-dimensional problem is solved on a
  separate processor. The trial results at the point \(x^k\) obtained for the problem being solved by
  particular processor are interpreted as the results of the trials in the rest problems (in the
  corresponding points \(x^{k_1},\dots,x^{k_S})\). In this approach, a trial at the point \(x^k \in
  [0,1]\) executed in the framework of the \(s\)-th problem, consists in the following sequence of
  operations:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Determine the image \(y^k=y^s (x^k)\) for the evolvent \(y^s (x)\).
    \item[Step 2.] Inform the rest of processors about the start of the trial execution at the point \( y^k\) (the
    blocking of the point \(y^k\) ).
    \item[Step 3.] Determine the preimages \(x{}^{k_s}  \in [0,1], 1\leqslant s\leqslant S\), of the point \(y^k\) and interpret the
    trial executed at the point \(y^k \in D \) as the execution of the trials in the \(S\) points
    \(x{}^{k_1} ,\dots,x{}^{k_s} \)
    \item[Step 4.] Inform the rest of processors about the trial results at the point \(y^k\).
  \end{enumerate}
\end{frame}

@@

\begin{frame}
  \frametitle{Evolvents comparison}
  \begin{figure}[ht]
    \hspace*{-0.9cm}
    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_opt_pt_op.pdf} }}
    \subfloat[\(r=5.0\)]{{\includegraphics[width=.55\textwidth]{gklsS2d_same_r_opt_pt_op.pdf} }} \hspace*{4cm}
    \caption{Operating characteristics on GKLS 2d Simple class}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Evolvents comparison}
  \begin{figure}[ht]
    \hspace*{-0.9cm}
    \subfloat[Minimal \(r\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_opt_pt_op.pdf} }}
    \subfloat[\(r=4.5\)]{{\includegraphics[width=.55\textwidth]{gklsS3d_same_r_opt_pt_op.pdf} }}
    \caption{Operating characteristics on GKLS 3d Simple class}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Choice of evolvent for the parallel algorithm}
  \begin{table}
  \begin{center}
  \caption{Averaged number of computations of \(g_0\) and of \(\varphi\) when solving the
  problems from GKLS 3d Simple class using the shifted evolvent}
    \begin{tabular}{|l|{c}|{c}|{c}|}
      \hline
    $L$ & $calc(g_0)$ & $calc(\varphi)$ & $\frac{calc(g_0)}{calc(\varphi)}$ ratio \\
    \hline
    2 & 96247.9  & 6840.14 & 14.07\\
    \hline
    3 & 153131.0 & 7702.82 & 19.88\\
    \hline
    \end{tabular}
    \label{tab:shifted_g0}
  \end{center}
  \end{table}
\end{frame}

@@

  \begin{table}
    \centering
    \caption{Speedup of parallel computations executed by the parallel algorithm}
    \label{tab:speedup}
    \begin{tabular}{cccccccc}
      \cline{3-8}\noalign{\smallskip}
      \multicolumn{2}{c}{  } & \textit{p} & \multicolumn{2}{c}{$N=4$} & &
  \multicolumn{2}{c}{$N=5$}   \\
      \noalign{\smallskip} \cline{4-5} \cline{7-8}  \noalign{\smallskip}
      \multicolumn{2}{c}{  } & & \textit{Simple} & \textit{Hard} & & \textit{Simple} &
  \textit{Hard}  \\
      \noalign{\smallskip}\hline
      I &
      \parbox{0.25\textwidth}{
      \begin{center}
      \textbf{1 cluster node}
      \end{center}		}
      & \textit{1}   & 12167(10.58s) & 25635(22.26s) & & 20979(22.78s) & 187353(205.83s)  \\
    &  & \textit{32} & 37.1(18.03) & 20.2(8.55)  & &  23.3(8.77) & 15.4(9.68) \\
    \hline \noalign{\smallskip}
  II  & \textbf{4 cluster nodes}  %\multirow{3}{*}{}
  & \textit{1} &        0.5(0.33) & 2.3(0.86)  & & 14.3(6.61) & 11.0(6.06) \\
  &   & \textit{32} & 190.1(9.59) & 28.1(1.08) & & 446.4(19.79) & 543.0(43.60) \\
    \noalign{\smallskip}\hline	\noalign{\smallskip}
  III & \textbf{8 cluster nodes} %\multirow{3}{*}{}
  & \textit{1}    & 15.0(6.05)  & 5.9(2.36)   & & 24.2(17.56)  & 32.9(24.87)  \\
  & & \textit{32} & 357.9(2.36) & 228.9(2.64) & & 582.8(20.96) & 793.0(33.89) \\
      \noalign{\smallskip}\hline
    \end{tabular}
  \end{table}

\begin{frame}
  \frametitle{Conclusions}
    %Already done:
    \begin{itemize}
      \item the smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
      \item the shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint. About 95\% of iterations are overhead to fight the auxiliary constraint.
      \item rotated evolvents perform almost the same as the shifted ones but without overhead.
      \item parallel optimization method shows up to 33x speedup on hard \(5d\) problems when using a set of rotated evolvents.
    \end{itemize}
    %Future work:
    %\begin{itemize}
    %  \item
    %\end{itemize}
\end{frame}

\begin{frame}{{}}
  \frametitle{Q\&A}
  \begin{center}
    \Large{Contacts:}
\vspace{0.5cm}
